Open-source tools in R for landscape ecology

### 2021-04-25 Comments_Reviewer.Rmd
Ups...forgot to knit
Adding synthax info to map section

**Reply**: We added an explanation how these package allow to incorporate functionality currently only available in GIS software into reproducible and shareable R scripts \vskip 0.5cm

@@

**Reply**: We agree that this is problematic and we weren't even aware this is happeing so thanks for pointing it out. We deactivated the automatic hyphenation by LaTeX. \vskip 0.5cm

@@

**Reply**: Thank you for checking the table so extensively. We made sure all packages should be included now. \textcolor{red}{Check again all packages are there} \vskip 0.5cm

@@

**Reply**: We apologise for this. The manuscript was written using the rticles packages with the corresponding template for Springer journals which somehow did not include journal names. We fixed the citation style including journal names. \textcolor{red}{STILL NEED TO DO THIS ACTUALLY!} \vskip 0.5cm

@@

**Reply**: We agree that the use of viridis might be confusing without mentioning it in the manuscript and we removed it from the R script.
We agree that ggplot2 graphics are a very large field.
This also highlights that it is outside the scope of this paper.
Nevertheless, ggplot2 is used quite frequently by the R community, so we decided to include it in the example.
We added a reference to the manuscript pointing towards ggplot2 specific resources. \vskip 0.5cm

>>>>

**Reply**: We added an explanation how these package allow to incorporate functionality currently only available in GIS software into reproducible and shareable R scripts. \vskip 0.5cm

@@

**Reply**: We agree that this is problematic and we weren't even aware this is happening so thanks for pointing it out. 
We deactivated the automatic hyphenation by LaTeX. \vskip 0.5cm

@@

**Reply**: Thank you for checking the table so extensively. 
We made sure all packages should be included now. We also added the information to the table caption that not all packages from the manuscript are included but only packages specifically designed for spatial data and landscape ecology. \vskip 0.5cm

@@

**Reply**: We apologize for this. 
The manuscript was written using the rticles packages with the corresponding template for Springer journals, which somehow did not include journal names. 
We fixed the citation style including journal names. \textcolor{red}{STILL NEED TO DO THIS ACTUALLY!} \vskip 0.5cm

@@

**Reply**: We agree that the use of viridis might be confusing without mentioning it in the manuscript and we removed it from the R script.
We agree that ggplot2 graphics are a very large field.
This also highlights that it is outside the scope of this paper.
Nevertheless, ggplot2 is used quite frequently by the R community, so we decided to include it in the example.
We added an explanation how base R maps and ggplot2 maps differ in their underlying synthax and capabilities. 
Also, we added a reference to the manuscript pointing towards ggplot2 specific resources. \vskip 0.5cm

### 2021-04-25 paper/paper.Rmd
Adding synthax info to map section

Creating maps is essential when working with spatial data.
Maps play an important role in checking the spatial and value-related quality of data, data exploration, and finally communicating results.
*R* allows two major types of maps.
Firstly, static maps in which the developer has full control over the presentation of the map and secondly, interactive maps in which the user has the possibility to modify the map by e.g. changing the displayed values.
All packages listed in Section \@ref(sec:spatrep) have build-in methods for plotting spatial data using the generic `plot()` function (Fig. \@ref(fig:fig-map)a).
However, the generic functions are focused on quick visual inspection of the data, rather than the creating complete maps.

Further plotting methods for raster objects can be found in the *rasterVis* package \cite{Lamigueiro2020}.
Also the popular plotting package *ggplot2* \cite{Wickham2016} (Fig. \@ref(fig:fig-map)b) has an extension especially designed for plotting spatial data named *ggspatial* \cite{Dunnington2020}. 
\textcolor{blue}{Fore more information about *ggplot2* in general see \cite{Wickham2016}.
Static thematic maps, including proportional symbols, choropleth, or typology maps, can be created with the *cartography* package \cite{Giraud2016}.
The *tmap* package provides a coherent plotting system for static and interactive maps that is based on the layered grammar of graphics \cite{Tennekes2018} and aims for creating publication-ready maps.
Quick interactive visualization of spatial data can be done with the *mapview* package\cite{Appelhans2020}.
Both, *tmap* (Fig. \@ref(fig:fig-map)c) and *mapview* build upon the *leaflet* package and leaflet javascript library \cite{Cheng2021}.
A slightly different approach to visualizing spatial data is adapted by the *rayshader* package \cite{Morgen-Wall2020} that creates topographic 2D and 3D maps.

(ref:capfigmap) Comparison of different options to create maps using a) base plot, b) *ggplot2* package, and c) *tmap* package. All maps show the total annual precipitation of Switzerland. Data were downloaded with the *raster* package (precipitation data) and *rnaturalearth* package (country borders) package. The code to create the maps can be found in the Appendix.
```{r fig-map, echo = FALSE, out.width="\\linewidth", fig.cap = "(ref:capfigmap)"}
include_graphics("data/Figure_2.pdf")
```

>>>>

Creating maps is essential when working with spatial data.
Maps play an important role in checking the spatial and value-related quality of data, data exploration, and finally communicating results.
*R* allows two major types of maps.
Firstly, static maps in which the developer has full control over the presentation of the map and secondly, interactive maps in which the user has the possibility to modify the map by e.g. changing the displayed values.
All packages listed in Section \@ref(sec:spatrep) have build-in methods for plotting spatial data using the generic `plot()` function (Fig. \@ref(fig:fig-map)a).
However, the generic functions are focused on quick visual inspection of the data (Fig. \@ref(fig:fig-map)a), rather than the creating complete maps \textcolor{blue}{as they do not directly support additional map elements (e.g., scale bar, north arrow) nor multiples maps.}

The popular plotting package *ggplot2* (Fig. \@ref(fig:fig-map)b) \textcolor{blue}{can also visualize spatial data and is build on a layered grammar of graphics, which allows to create maps by combining individual graphical elements (e.g., raster and vector elements) \cite{Wickham2016}.
Additionally, there is an extension especially designed to plot spatial data named *ggspatial*, which provides additional map elements \cite{Dunnington2020}.
For more information about *ggplot2* in general, see \cite{Wickham2016}}.
The *tmap* package provides a provides a coherent plotting system intended for drawing maps (Fig. \@ref(fig:fig-map)c) \cite{Tennekes2018}. 
\textcolor{blue}{This package operates in two modes, static and interactive, which means that the same code can be used to create static or interactive maps.}
Quick interactive visualization of spatial data can be done with the *mapview* package\cite{Appelhans2020}.
Both, *tmap* and *mapview* build upon the *leaflet* package and leaflet javascript library \cite{Cheng2021}.
Static thematic maps, including proportional symbols, choropleth, or typology maps, can be created with the *cartography* package \cite{Giraud2016}.
Further plotting methods for raster objects can be found in the *rasterVis* package \cite{Lamigueiro2020}.
A slightly different approach to visualizing spatial data is adapted by the *rayshader* package \cite{Morgen-Wall2020} that creates topographic 2D and 3D maps.

(ref:capfigmap) Comparison of different options to create maps of the total annual precipitation in Switzerland using a) base plot, b) *ggplot2* package, and c) *tmap* package. All three approaches result in similar-looking outputs, but their syntax and capabilities differ. All maps show the total annual precipitation of Switzerland. Data were downloaded with the *raster* package (precipitation data) and *rnaturalearth* package (country borders) package. The code to create the maps can be found in the Appendix.
```{r fig-map, echo = FALSE, out.width="\\linewidth", fig.cap = "(ref:capfigmap)"}
include_graphics("data/Figure_2.pdf")
```

### 2021-04-25 paper/paper.tex
Adding synthax info to map section

Creating maps is essential when working with spatial data.
Maps play an important role in checking the spatial and value-related quality of data, data exploration, and finally communicating results.
\emph{R} allows two major types of maps.
Firstly, static maps in which the developer has full control over the presentation of the map and secondly, interactive maps in which the user has the possibility to modify the map by e.g.~changing the displayed values.
All packages listed in Section \ref{sec:spatrep} have build-in methods for plotting spatial data using the generic \texttt{plot()} function (Fig. \ref{fig:fig-map}a).
However, the generic functions are focused on quick visual inspection of the data, rather than the creating complete maps.

Further plotting methods for raster objects can be found in the \emph{rasterVis} package \cite{Lamigueiro2020}.
Also the popular plotting package \emph{ggplot2} \cite{Wickham2016} (Fig. \ref{fig:fig-map}b) has an extension especially designed for plotting spatial data named \emph{ggspatial} \cite{Dunnington2020}.
\textbackslash textcolor\{blue\}\{Fore more information about \emph{ggplot2} in general see \cite{Wickham2016}.
Static thematic maps, including proportional symbols, choropleth, or typology maps, can be created with the \emph{cartography} package \cite{Giraud2016}.
The \emph{tmap} package provides a coherent plotting system for static and interactive maps that is based on the layered grammar of graphics \cite{Tennekes2018} and aims for creating publication-ready maps.
Quick interactive visualization of spatial data can be done with the \emph{mapview} package\cite{Appelhans2020}.
Both, \emph{tmap} (Fig. \ref{fig:fig-map}c) and \emph{mapview} build upon the \emph{leaflet} package and leaflet javascript library \cite{Cheng2021}.
A slightly different approach to visualizing spatial data is adapted by the \emph{rayshader} package \cite{Morgen-Wall2020} that creates topographic 2D and 3D maps.

\begin{figure}
\includegraphics[width=\linewidth]{data/Figure_2} \caption{Comparison of different options to create maps using a) base plot, b) \emph{ggplot2} package, and c) \emph{tmap} package. All maps show the total annual precipitation of Switzerland. Data were downloaded with the \emph{raster} package (precipitation data) and \emph{rnaturalearth} package (country borders) package. The code to create the maps can be found in the Appendix.}\label{fig:fig-map}
\end{figure}

>>>>

Creating maps is essential when working with spatial data.
Maps play an important role in checking the spatial and value-related quality of data, data exploration, and finally communicating results.
\emph{R} allows two major types of maps.
Firstly, static maps in which the developer has full control over the presentation of the map and secondly, interactive maps in which the user has the possibility to modify the map by e.g.~changing the displayed values.
All packages listed in Section \ref{sec:spatrep} have build-in methods for plotting spatial data using the generic \texttt{plot()} function (Fig. \ref{fig:fig-map}a).
However, the generic functions are focused on quick visual inspection of the data (Fig. \ref{fig:fig-map}a), rather than the creating complete maps \textcolor{blue}{as they do not directly support additional map elements (e.g., scale bar, north arrow) nor multiples maps.}

The popular plotting package \emph{ggplot2} (Fig. \ref{fig:fig-map}b) \textcolor{blue}{can also visualize spatial data and is build on a layered grammar of graphics, which allows to create maps by combining individual graphical elements (e.g., raster and vector elements) \cite{Wickham2016}.
Additionally, there is an extension especially designed to plot spatial data named *ggspatial*, which provides additional map elements \cite{Dunnington2020}.
For more information about *ggplot2* in general, see \cite{Wickham2016}}.
The \emph{tmap} package provides a provides a coherent plotting system intended for drawing maps (Fig. \ref{fig:fig-map}c) \cite{Tennekes2018}.
\textcolor{blue}{This package operates in two modes, static and interactive, which means that the same code can be used to create static or interactive maps.}
Quick interactive visualization of spatial data can be done with the \emph{mapview} package\cite{Appelhans2020}.
Both, \emph{tmap} and \emph{mapview} build upon the \emph{leaflet} package and leaflet javascript library \cite{Cheng2021}.
Static thematic maps, including proportional symbols, choropleth, or typology maps, can be created with the \emph{cartography} package \cite{Giraud2016}.
Further plotting methods for raster objects can be found in the \emph{rasterVis} package \cite{Lamigueiro2020}.
A slightly different approach to visualizing spatial data is adapted by the \emph{rayshader} package \cite{Morgen-Wall2020} that creates topographic 2D and 3D maps.

\begin{figure}
\includegraphics[width=\linewidth]{data/Figure_2} \caption{Comparison of different options to create maps of the total annual precipitation in Switzerland using a) base plot, b) \emph{ggplot2} package, and c) \emph{tmap} package. All three approaches result in similar-looking outputs, but their syntax and capabilities differ. All maps show the total annual precipitation of Switzerland. Data were downloaded with the \emph{raster} package (precipitation data) and \emph{rnaturalearth} package (country borders) package. The code to create the maps can be found in the Appendix.}\label{fig:fig-map}
\end{figure}
